---
title: "EECSE6690 - Final Project Code"
output: html_notebook
---

Author: Kosta Andoni (ka2604)
Date: 05/10/2018

This RMarkdown notebook goes through the process of steps taken to reproduce the results from the Sapsanis paper and then continue with my own work with the SVM. Each R code chunk performs a different task in the whole process. The code chunks are meant to be run in order to get the results. For the most part, code chunks are standalone if all the prerequisite data and packages are loaded.

```{r}
#Option to suppress warnings in output
options(warn=-1)
```


First, we load the EMG data from all the subjects. To reproduce the paper's results, each subjects runs must be stored separately. We can store the data together in intermediary steps as long as we keep track of which samples belong to which subject. There are three female subjects and two male subjects, but gender isn't an important variable in this study.

```{r}
#Using the R.matlab package, we read the data provided in the dataset from the UCI Data Repository at the following link: https://archive.ics.uci.edu/ml/datasets/sEMG+for+Basic+Hand+movements#
#In order for the code to run properly, the data matrices must be placed into a directory named "Database 1" at the same directory level as this RMD file.

library(R.matlab)
f1 = readMat("Database 1/female_1.mat")
f2 = readMat("Database 1/female_2.mat")
f3 = readMat("Database 1/female_3.mat")
m1 = readMat("Database 1/male_1.mat")
m2 = readMat("Database 1/male_2.mat")
```


Lets rearrange the data into a few matrices and lists to make our analysis easier as we move forward and to give our variables meaningful names.

```{r}
#Row index in all matrices/vectors correspond
EMG.raw = matrix(data=NA,1800,3000)       #Stores the raw EMG data
subj.list = rep(list(bquote()),1800)      #Stores the subject number
grasp.list = rep(list(bquote()),1800)     #Stores the hand movement/grasp
ch.list = rep(list(bquote()),1800)        #Stores the channel number
rep.list = rep(list(bquote()),1800)       #Stores the repitition number

#Create a list of all our subjects
subject.list = c("f1","f2","f3","m1","m2")
subject.num = 0
seq.num = 0

#Iterate through the list of subjects, each grasp, and each sequence to extract the raw info
for (subject in subject.list) {
  subject.num = subject.num + 1
  for (grasp in names(get(subject))) {
    grasp.type = sub("\\..*$", "", grasp)         #Parse the grasp name
    ch.num = sub('.*(?=.$)', '', grasp, perl=T)   #Parse the channel number
    for (row in seq(1,30)) {
      seq.num = seq.num + 1
      subj.list[seq.num] = subject.num
      grasp.list[seq.num] = grasp.type
      ch.list[seq.num] = ch.num
      rep.list[seq.num] = row
      EMG.raw[seq.num,] = eval(parse(text = paste(subject,"$",grasp,sep="")))[row,]
    
    }
  }
}

```


Exclude the pre-activation portions of each grasp to only keep the action. This is done through a sliding window approach that averages the raw EMG across the window. If the averaged EMG value crosses a certain threshold, we take the action to start at this point and discard all previous samples.

```{r}
sliding.window = 20       #Set the window to 40 ms

#Function: slidingIEMG
#Input: EMG.x - Data vector of raw EMG time series
#       sliding.window - Number of samples in the window
#Output: IEMG - Data vector of EMG data averaged in the window
#Description: A sliding window approach to average data. IEMG stands for 
#             Integrated Electromyogram

slidingIEMG = function(EMG.x, sliding.window) {
  #Initialize
  window.start=1
  i = 1
  IEMG = vector(length=floor(length(EMG.x)/sliding.window))
  
  #While the end of the window doesn't exceed the bounds of the data vector,
  #  we can slide the window and take the average of the samples that fall in it
  while( (window.start + sliding.window - 1) <= length(EMG.x) ) {
    window.end = window.start + sliding.window - 1        #Calculate end of window
    IEMG[i] = sum(abs(EMG.x[window.start:window.end]))    #Averaging & storing
    window.start = window.start + 1                       #Slide the window
    i = i + 1
  }
  
  #Return the final data vector with the averaged EMG
  return(IEMG)
}

#Function: trimEMG
#Input: EMG.x - Data vector containing raw EMG time series
#       IEMG - Data vector of averaged EMG data via sliding window
#Output: A trimmed data vector with inactive samples removed
#Description: Trims the inactive regions/samples of the data based on when the IEMG exceeds 
#             the threshold

trimEMG = function(EMG.x,IEMG) {
  threshold = (max(IEMG)-min(IEMG))/2         #Threshold is half the range
  cutoff = min(which(IEMG > threshold)) + 1   #Cut the data where the threshold is crossed
  return(EMG.x[-(1:cutoff-1)])                #Return the cut vector
}

#__ Main Code __

#Initialize the vector to store the trimmed representations
EMG.trim = matrix(data=NA,1800,3000)

#For each EMG sequence, trim away the inactive regions - effectively offsetting the start of the sequence to the start of the hand movement
for(seq in seq(1:nrow(EMG.raw))) {
  IEMG = slidingIEMG(EMG.raw[seq,],sliding.window)    #Calculate the sliding IEMG vector
  seq.trimmed = trimEMG(EMG.raw[seq,],IEMG)           #Trim the sequence
  length(seq.trimmed)                                 #Get the new number of samples in sequence
  #Store the trimmed sequence and buffer the end of the row with NA
  EMG.trim[seq,] = c(seq.trimmed, rep(NA, 3000 - length(seq.trimmed)))
}


```


Define a data table constructor

```{r}
#Function: createDataTable
#Input: NA
#Output: EMG.df - A data table that stores all the features of interest in our use case
#Description: Generates and empty data table with columns for each of our features of interest.
#             A data table is used rather than a data frame so we can store lists in a column (i.e. 3D)

createDataTable = function(){
  EMG.df=setNames(data.table(matrix(ncol = 56, nrow = 0)),
        c("subj", "grasp", "ch", "rep", "segNum", "segment",
          "emgIEMG", "emgZeroCross", "emgSSC", "emgWL", "emgWAMP", "emgVAR", "emgSkew", "emgKurt", 
          "imf1IEMG", "imf1ZeroCross", "imf1SSC", "imf1WL", "imf1WAMP", "imf1VAR", "imf1Skew", 
          "imf1Kurt", "imf1IFMedian", "imf1IFSD", "imf1IFKurt", 
          "imf2IEMG", "imf2ZeroCross", "imf2SSC", "imf2WL", "imf2WAMP", "imf2VAR", "imf2Skew", 
          "imf2Kurt", "imf2IFMedian", "imf2IFSD", "imf2IFKurt", 
          "imf3IEMG", "imf3ZeroCross", "imf3SSC", "imf3WL", "imf3WAMP", "imf3VAR", "imf3Skew", 
          "imf3Kurt", "imf3IFMedian", "imf3IFSD", "imf3IFKurt", 
          "resIEMG", "resZeroCross", "resSSC", "resWL", "resWAMP", "resVAR", "resSkew", "resKurt",
          "featVec"))
  
  return(EMG.df)
}
```


Segment the trimmed EMG signals into 300ms segments (overlapping window of 30ms)

```{r}
library(data.table)
library(EMD)

#Function: bindDataTable
#Input: df - Existing EMG data table
#       subject.num, grasp.type, ch.num, rep.num, seg.num, segment - Data for new row of data table
#Output: df - EMG data table with new row added
#Description: Binds a new row of data to the data table

bindDataTable = function(df, subject.num, grasp.type, ch.num, rep.num, seg.num, segment) {
  df = rbindlist(list(df, data.table(subj=subject.num, grasp=grasp.type, ch=ch.num, rep=rep.num, 
                                segNum = seg.num, segment=list(segment))), fill=T, use.names=T)
  return(df)
}

#Function: segmentSequence
#Input: EMG.df - EMG data table to store segments and their resective stats
#       EMG.x - Data vector containing EMG time series for a sequence
#       EMG.grasp - Type of grasp
#       EMG.ch - EMG channel data was recorded on
#       EMG.rep - Repition number of movement
#       length - Segment length
#       stride - Distance between segment starts
#Output: EMG.df - EMG data table with the segments of the input sequence added
#Description: Partitions an input sequence into segments and appends each segment to a data table

segmentSequence = function(EMG.df, EMG.x, EMG.subj, EMG.grasp, EMG.ch, EMG.rep, length, stride) {
  #Initialize parameter
  window.start = 1
  window.end = window.start + length - 1
  i = 1

  #While there are still segments to extract in the sequence,
  #Note: Be careful not to go out of bounds or include any NAs in the segment
  while(((window.end) <= length(EMG.x)) && (!is.na(EMG.x[window.end]))) {
    
    #Grab a segment
    segment = EMG.x[window.start:window.end]
    
    #Append it to the data table
    EMG.df = bindDataTable(EMG.df, EMG.subj, EMG.grasp, EMG.ch, EMG.rep, i, segment)
    
    #Update parameters
    window.start = window.start + stride
    window.end = window.start + length - 1
    i = i +1
  }
  
  #Return populated data table
  return(EMG.df)
}

#__ Main Code __

#Initialize the master data table for segments
EMG.df = createDataTable()

#Initialize parameter
segment.length = 150
segment.stride = 15

#For each sequence available, ...
for(seq in seq(1:nrow(EMG.trim))) {
  
  #Bind to master data table in sequence groups of 10 rather than every seqeunce
  #(This is a performance optimization)
  if(seq%%10 == 1){
    if(seq > 1) {
      EMG.df = rbindlist(list(EMG.df,EMG.df.tmp))
    }
    EMG.df.tmp = createDataTable()
  }
  
  #Segment the sequence into segments
  EMG.df.tmp = segmentSequence(EMG.df.tmp, EMG.trim[seq,], subj.list[seq], grasp.list[seq],
                               ch.list[seq], rep.list[seq], segment.length, segment.stride)
}

```


Calculate the first three IMFs using EMD for 300ms windows, 30ms apart. Apply the 8 features as listed below

```{r}
library(EMD)
library(e1071)
library(foreach)
library(doParallel)

#Function: updateDataTableEMD
#Input: df - Data table row corresponding to a segment
#Output: df - Updated data table with important statistic
#Description: Calculates all the statistics for each segment. Warning: Takes hours to run
#             even with parallelization

updateDataTableEMD = function(df){
  
  seg.x = unlist(df$segment)        #Store the segment data as a simple vector
  decomp.x = emd(seg.x, max.imf=3)  #Decompose the segment data into 3 IMFs and a residual
                                    #  using the EMD algorithm
  
  #Apply the Hilbert transform to each IMF to get the IF
  decomp.if = hilbertspec(decomp.x$imf)
  
  #Calculate the IEMG
  df[, emgIEMG := as.numeric(mean(abs(seg.x)))]
  df[, imf1IEMG := as.numeric(mean(abs(decomp.x$imf[,1])))]
  df[, imf2IEMG := as.numeric(mean(abs(decomp.x$imf[,2])))]
  df[, imf3IEMG := as.numeric(mean(abs(decomp.x$imf[,3])))]
  df[, resIEMG := as.numeric(mean(abs(decomp.x$residue)))]
  
  #Calculate the zero crossing
  df[, emgZeroCross := as.numeric(sum(abs(diff(sign(seg.x))))/2)]
  df[, imf1ZeroCross := as.numeric(sum(abs(diff(sign(decomp.x$imf[,1]))))/2)]
  df[, imf2ZeroCross := as.numeric(sum(abs(diff(sign(decomp.x$imf[,2]))))/2)]
  df[, imf3ZeroCross := as.numeric(sum(abs(diff(sign(decomp.x$imf[,3]))))/2)]
  df[, resZeroCross := as.numeric(sum(abs(diff(sign(decomp.x$residue))))/2)]

  #Calculate the slope sign change
  df[, emgSSC := as.numeric(sum(abs(diff(sign(diff(seg.x))))))/2]
  df[, imf1SSC := as.numeric(sum(abs(diff(sign(diff(decomp.x$imf[,1]))))))/2]
  df[, imf2SSC := as.numeric(sum(abs(diff(sign(diff(decomp.x$imf[,2]))))))/2]
  df[, imf3SSC := as.numeric(sum(abs(diff(sign(diff(decomp.x$imf[,3]))))))/2]
  df[, resSSC := as.numeric(sum(abs(diff(sign(diff(decomp.x$residue))))))/2]
  
  #Calculate the Willison Amplitude
  df[, emgWL := as.numeric(sum(abs(diff(seg.x))))]
  df[, imf1WL := as.numeric(sum(abs(diff(decomp.x$imf[,1]))))]
  df[, imf2WL := as.numeric(sum(abs(diff(decomp.x$imf[,2]))))]
  df[, imf3WL := as.numeric(sum(abs(diff(decomp.x$imf[,3]))))]
  df[, resWL := as.numeric(sum(abs(diff(decomp.x$residue))))]
  
  #Calculate the waveform length
  df[, emgWAMP := as.numeric(length(which(abs(diff(seg.x)) > 1.5)))]
  df[, imf1WAMP := as.numeric(length(which(abs(diff(decomp.x$imf[,1])) > 1.0)))]
  df[, imf2WAMP := as.numeric(length(which(abs(diff(decomp.x$imf[,2])) > 0.5)))]
  df[, imf3WAMP := as.numeric(length(which(abs(diff(decomp.x$imf[,3])) > 0.25)))]
  df[, resWAMP := as.numeric(length(which(abs(diff(decomp.x$residue)) > 0.1)))]
  
  #Calculate the Variance
  df[, emgVAR := as.numeric(var(seg.x))]
  df[, imf1VAR := as.numeric(var(decomp.x$imf[,1]))]
  df[, imf2VAR := as.numeric(var(decomp.x$imf[,2]))]
  df[, imf3VAR := as.numeric(var(decomp.x$imf[,3]))]
  df[, resVAR := as.numeric(var(decomp.x$residue))]
  
  #Calculate the Skewness
  df[, emgSkew := as.numeric(skewness(seg.x))]
  df[, imf1Skew := as.numeric(skewness(decomp.x$imf[,1]))]
  df[, imf2Skew := as.numeric(skewness(decomp.x$imf[,2]))]
  df[, imf3Skew := as.numeric(skewness(decomp.x$imf[,3]))]
  df[, resSkew := as.numeric(skewness(decomp.x$residue))]
  
  #Calculate the Kurtosis
  df[, emgKurt := as.numeric(kurtosis(seg.x))]
  df[, imf1Kurt := as.numeric(kurtosis(decomp.x$imf[,1]))]
  df[, imf2Kurt := as.numeric(kurtosis(decomp.x$imf[,2]))]
  df[, imf3Kurt := as.numeric(kurtosis(decomp.x$imf[,3]))]
  df[, resKurt := as.numeric(kurtosis(decomp.x$residue))]
  
  #Calculate the Median of the IFs
  df[, imf1IFMedian := as.numeric(median(decomp.if$instantfreq[,1]))]
  df[, imf2IFMedian := as.numeric(median(decomp.if$instantfreq[,2]))]
  df[, imf3IFMedian := as.numeric(median(decomp.if$instantfreq[,3]))]
  
  #Calculate the Standard Deviation of the IFs
  df[, imf1IFSD := as.numeric(sd(decomp.if$instantfreq[,1]))]
  df[, imf2IFSD := as.numeric(sd(decomp.if$instantfreq[,2]))]
  df[, imf3IFSD := as.numeric(sd(decomp.if$instantfreq[,3]))]
  
  #Calculate the Kurtosis of the IFs
  df[, imf1IFKurt := as.numeric(kurtosis(decomp.if$instantfreq[,1]))]
  df[, imf2IFKurt := as.numeric(kurtosis(decomp.if$instantfreq[,2]))]
  df[, imf3IFKurt := as.numeric(kurtosis(decomp.if$instantfreq[,3]))]
  
  return(df)
}

#__ Main Code__

#Initialize the data tables
EMG.df.update = createDataTable()
EMG.df.tmp = createDataTable()

#Configure the core cluster to perform the parallel computing
cores=detectCores()
cl <- makeCluster(cores[1]-1, useXDR=F)
registerDoParallel(cl)

#Implement updating the EMG data table features using parallel programming
EMG.df.update <- foreach(i=1:nrow(EMG.df), .combine=function(x,y)rbindlist(list(x,y)), 
                         .packages=c('EMD','data.table','e1071'), .verbose=T) %dopar% {
  
  #Update the data table for each segment                         
  EMG.df.tmp = updateDataTableEMD(EMG.df[i])

  #Bind the temp data.table to the updated EMG data table
  EMG.df.tmp 
}

#Release the core cluster
stopCluster(cl)

#Replace the master data table with the updated statistics
EMG.df = copy(EMG.df.update)

```


Trim the data table into a data frame. Also, factor the hand movements into 1 through 6.
1: Spherical - spher 
2: Tip - tip
3: Palmar - palm
4: Lateral - lat
5: Cylindrical - cyl
6: Hook - hook

```{r}
#Create a data table for just the features - leaving the EMG data behind
EMG.df.features = copy(EMG.df)
EMG.df.features[, c("featVec", "segment") := NULL]

#Convert types to numeric rather than list
EMG.df.features[, subj := as.numeric(EMG.df.features[, subj])]
EMG.df.features[, ch := as.numeric(EMG.df.features[, ch])]
EMG.df.features[, rep := as.numeric(EMG.df.features[, rep])]

#Factor the grasp data
EMG.df.features[, grasp := factor(EMG.df.features[, grasp], 
                                  levels=c('spher','tip','palm','lat','cyl','hook'),
                                  labels=c('S','T','P','L','C','H'))]

#Convert data table to data frame                                 
setDF(EMG.df.features)
```


Perform PCA to remove redundancies in the feature vector. Not implemented further than dimension reduction investigation. Data is not reduced for subsequent steps but this is where we would start.

```{r}
#Do PCA - without the qualitative variables
EMG.pca = prcomp(EMG.df.features[, !names(EMG.df.features) %in%
                                   c("subj","ch","rep","segNum","grasp")], scale. = T, tol=0.05)
summary(EMG.pca)

```


Divide the dataset per subject - including training and testing sets. Put the feature vector through the linear classifier - LDA approach - and present the confusion matrix

```{r}
library(MASS)
library(caret)

#Initialize the seeds for 5x2 cross validation
seeds = c(04081994,07041994,06122015,09202015,05102018)

#Initialize vectors to store the accumulated predictions across 5 loops
EMG.subj1.pred.accum = character()
EMG.subj2.pred.accum = character()
EMG.subj3.pred.accum = character()
EMG.subj4.pred.accum = character()
EMG.subj5.pred.accum = character()

#Initialize vectors to store the accumulated actual values across 5 loops
EMG.subj1.test.accum = character()
EMG.subj2.test.accum = character()
EMG.subj3.test.accum = character()
EMG.subj4.test.accum = character()
EMG.subj5.test.accum = character()

#Iterate through each seed to implement 5x2 cross validation
for( i in seq(1,length(seeds))) {
  set.seed(seeds[i])
  
  #For each subject, pull their data from the master data frame - EMG.df.features
  #Then, split the data into a training and testing set (50/50)
  
  #Subject 1
  EMG.df.subj1 = subset(EMG.df.features, subj==1)
  
  sample_size = floor(0.5 * nrow(EMG.df.subj1))
  train_indices = sample(seq_len(nrow(EMG.df.subj1)), size = sample_size)
  EMG.df.subj1.train = EMG.df.subj1[train_indices,]
  EMG.df.subj1.test = EMG.df.subj1[-train_indices,]
  
  #Subject2
  EMG.df.subj2 = subset(EMG.df.features, subj==2)
  
  sample_size = floor(0.5 * nrow(EMG.df.subj2))
  train_indices = sample(seq_len(nrow(EMG.df.subj2)), size = sample_size)
  EMG.df.subj2.train = EMG.df.subj2[train_indices,]
  EMG.df.subj2.test = EMG.df.subj2[-train_indices,]
  
  #Subject3
  EMG.df.subj3 = subset(EMG.df.features, subj==3)
  
  sample_size = floor(0.5 * nrow(EMG.df.subj3))
  train_indices = sample(seq_len(nrow(EMG.df.subj3)), size = sample_size)
  EMG.df.subj3.train = EMG.df.subj3[train_indices,]
  EMG.df.subj3.test = EMG.df.subj3[-train_indices,]
  
  #Subject4
  EMG.df.subj4 = subset(EMG.df.features, subj==4)
  
  sample_size = floor(0.5 * nrow(EMG.df.subj4))
  train_indices = sample(seq_len(nrow(EMG.df.subj4)), size = sample_size)
  EMG.df.subj4.train = EMG.df.subj4[train_indices,]
  EMG.df.subj4.test = EMG.df.subj4[-train_indices,]
  
  #Subject5
  EMG.df.subj5 = subset(EMG.df.features, subj==5)
  
  sample_size = floor(0.5 * nrow(EMG.df.subj5))
  train_indices = sample(seq_len(nrow(EMG.df.subj5)), size = sample_size)
  EMG.df.subj5.train = EMG.df.subj5[train_indices,]
  EMG.df.subj5.test = EMG.df.subj5[-train_indices,]
  
  
  # Subject1 - Fit model
  EMG.subj1.lda = lda(grasp~. - subj - ch - rep - segNum, data=EMG.df.subj1.train)
  # Subject1 - Make predictions
  EMG.subj1.pred = predict(EMG.subj1.lda, EMG.df.subj1.test)$class
  # Subject1 - Append results to accumulation vectors
  EMG.subj1.pred.accum = c(EMG.subj1.pred.accum, EMG.subj1.pred)
  EMG.subj1.test.accum = c(EMG.subj1.test.accum, EMG.df.subj1.test$grasp)
  
  
  # Subject2 - Fit model
  EMG.subj2.lda = lda(grasp~. - subj - ch - rep - segNum, data=EMG.df.subj2.train)
  # Subject2 - Make predictions
  EMG.subj2.pred = predict(EMG.subj2.lda, EMG.df.subj2.test)$class
  # Subject2 - Append results to accumulation vectors
  EMG.subj2.pred.accum = c(EMG.subj2.pred.accum, EMG.subj2.pred)
  EMG.subj2.test.accum = c(EMG.subj2.test.accum, EMG.df.subj2.test$grasp)
  
  
  # Subject3 - Fit model
  EMG.subj3.lda = lda(grasp~. - subj - ch - rep - segNum, data=EMG.df.subj3.train)
  # Subject3 - Make predictions
  EMG.subj3.pred = predict(EMG.subj3.lda, EMG.df.subj3.test)$class
  # Subject3 - Append results to accumulation vectors
  EMG.subj3.pred.accum = c(EMG.subj3.pred.accum, EMG.subj3.pred)
  EMG.subj3.test.accum = c(EMG.subj3.test.accum, EMG.df.subj3.test$grasp)
  
  
  # Subject4 - Fit model
  EMG.subj4.lda = lda(grasp~. - subj - ch - rep - segNum, data=EMG.df.subj4.train)
  # Subject4 - Make predictions
  EMG.subj4.pred = predict(EMG.subj4.lda, EMG.df.subj4.test)$class
  # Subject4 - Append results to accumulation vectors
  EMG.subj4.pred.accum = c(EMG.subj4.pred.accum, EMG.subj4.pred)
  EMG.subj4.test.accum = c(EMG.subj4.test.accum, EMG.df.subj4.test$grasp)
  
  
  # Subject5 - Fit model
  EMG.subj5.lda = lda(grasp~. - subj - ch - rep - segNum, data=EMG.df.subj5.train)
  # Subject5 - Make predictions
  EMG.subj5.pred = predict(EMG.subj5.lda, EMG.df.subj5.test)$class
  # Subject5 - Append results to accumulation vectors
  EMG.subj5.pred.accum = c(EMG.subj5.pred.accum, EMG.subj5.pred)
  EMG.subj5.test.accum = c(EMG.subj5.test.accum, EMG.df.subj5.test$grasp)
  
}


#Factor all the prediction accumulation vectors for the confusion matrix
EMG.subj1.pred.accum = factor(EMG.subj1.pred.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj2.pred.accum = factor(EMG.subj2.pred.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj3.pred.accum = factor(EMG.subj3.pred.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj4.pred.accum = factor(EMG.subj4.pred.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj5.pred.accum = factor(EMG.subj5.pred.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))


#Factor all the test accumulation vectors for the confusion matrix
EMG.subj1.test.accum = factor(EMG.subj1.test.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj2.test.accum = factor(EMG.subj2.test.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj3.test.accum = factor(EMG.subj3.test.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj4.test.accum = factor(EMG.subj4.test.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))
EMG.subj5.test.accum = factor(EMG.subj5.test.accum, 
                              levels=1:6,
                              labels=c('S','T','P','L','C','H'))


#Create a confusion matrix for each subject's results
confusionMatrix(EMG.subj1.pred.accum, EMG.subj1.test.accum)
confusionMatrix(EMG.subj2.pred.accum, EMG.subj2.test.accum)
confusionMatrix(EMG.subj3.pred.accum, EMG.subj3.test.accum)
confusionMatrix(EMG.subj4.pred.accum, EMG.subj4.test.accum)
confusionMatrix(EMG.subj5.pred.accum, EMG.subj5.test.accum)

```


Now put the feature vector through SVM (with a radial kernel) and present the confusion matrix

```{r}
#Initialize 
cost = c(0.01, 0.1, 1, 10, 100)   #List of cost parameters to test
gamma= c(0.01, 0.1, 1, 10, 100)   #List of gamma parameters to test
samples.to.use=5000               #Limit of segments to use to make computation faster/less complex

#Scale the training data to normalize for SVM. Works but SVM quality is very poor.Needs to be explored.
#EMG.df.subj1.train[, !names(EMG.df.subj1.train) %in% c("subj","ch","rep","segNum","grasp")] = 
  #scale(EMG.df.subj1.train[, !names(EMG.df.subj1.train) %in% c("subj","ch","rep","segNum","grasp")])

#For each subject, tune the svm to find the lowest training error. 
#Then, use the best model to make predictions on the testing set
#Finally, present the confusion matrix

#Subject 1
EMG.subj1.svm.rbf = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj1.train[1:samples.to.use,],
                     kernel = "radial", ranges = list(cost=cost, gamma=gamma))
EMG.subj1.svm.pred.rbf = predict(EMG.subj1.svm.rbf$best.model, EMG.df.subj1.test)
confusionMatrix(EMG.subj1.svm.pred.rbf, EMG.df.subj1.test$grasp)

#Subject 2
EMG.subj2.svm.rbf = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj2.train[1:samples.to.use,],
                     kernel = "radial", ranges = list(cost=cost, gamma=gamma))
EMG.subj2.svm.pred.rbf = predict(EMG.subj2.svm.rbf$best.model, EMG.df.subj2.test)
confusionMatrix(EMG.subj2.svm.pred.rbf, EMG.df.subj2.test$grasp)

#Subject 3
EMG.subj3.svm.rbf = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj3.train[1:samples.to.use,],
                     kernel = "radial", ranges = list(cost=cost, gamma=gamma))
EMG.subj3.svm.pred.rbf = predict(EMG.subj3.svm.rbf$best.model, EMG.df.subj3.test)
confusionMatrix(EMG.subj3.svm.pred.rbf, EMG.df.subj3.test$grasp)

#Subject 4
EMG.subj4.svm.rbf = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj4.train[1:samples.to.use,],
                     kernel = "radial", ranges = list(cost=cost, gamma=gamma))
EMG.subj4.svm.pred.rbf = predict(EMG.subj4.svm.rbf$best.model, EMG.df.subj4.test)
confusionMatrix(EMG.subj4.svm.pred.rbf, EMG.df.subj4.test$grasp)

#Subject 5
EMG.subj5.svm.rbf = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj5.train[1:samples.to.use,],
                     kernel = "radial", ranges = list(cost=cost, gamma=gamma))
EMG.subj5.svm.pred.rbf = predict(EMG.subj5.svm.rbf$best.model, EMG.df.subj5.test)
confusionMatrix(EMG.subj5.svm.pred.rbf, EMG.df.subj5.test$grasp)

```


Now put the feature vector through SVM (with a linear kernel) and present the confusion matrix

```{r}
#Initialize 
cost = c(0.01, 0.1, 1, 10, 100)   #List of cost parameters to test
samples.to.use=1000               #Limit of segments to use to make computation faster/less complex


#For each subject, tune the svm to find the lowest training error. 
#Then, use the best model to make predictions on the testing set
#Finally, present the confusion matrix

#Subject 1
EMG.subj1.svm.lin = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj1.train[1:samples.to.use,],
                     kernel = "linear", ranges = list(cost=cost, gamma=gamma))
EMG.subj1.svm.pred.lin = predict(EMG.subj1.svm.lin$best.model, EMG.df.subj1.test)
confusionMatrix(EMG.subj1.svm.pred.lin, EMG.df.subj1.test$grasp)

#Subject 2
EMG.subj2.svm.lin = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj2.train[1:samples.to.use,],
                     kernel = "linear", ranges = list(cost=cost, gamma=gamma))
EMG.subj2.svm.pred.lin = predict(EMG.subj2.svm.lin$best.model, EMG.df.subj2.test)
confusionMatrix(EMG.subj2.svm.pred.lin, EMG.df.subj2.test$grasp)

#Subject 3
EMG.subj3.svm.lin = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj3.train[1:samples.to.use,],
                     kernel = "linear", ranges = list(cost=cost, gamma=gamma))
EMG.subj3.svm.pred.lin = predict(EMG.subj3.svm.lin$best.model, EMG.df.subj3.test)
confusionMatrix(EMG.subj3.svm.pred.lin, EMG.df.subj3.test$grasp)

#Subject 4
EMG.subj4.svm.lin = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj4.train[1:samples.to.use,],
                     kernel = "linear", ranges = list(cost=cost, gamma=gamma))
EMG.subj4.svm.pred.lin = predict(EMG.subj4.svm.lin$best.model, EMG.df.subj4.test)
confusionMatrix(EMG.subj4.svm.pred.lin, EMG.df.subj4.test$grasp)

#Subject 5
EMG.subj5.svm.lin = tune(svm, grasp~. - subj - ch - rep - segNum,  
                     data=EMG.df.subj5.train[1:samples.to.use,],
                     kernel = "linear", ranges = list(cost=cost, gamma=gamma))
EMG.subj5.svm.pred.lin = predict(EMG.subj5.svm.lin$best.model, EMG.df.subj5.test)
confusionMatrix(EMG.subj5.svm.pred.lin, EMG.df.subj5.test$grasp)
```
